Supplementary Information

### 2021-08-04 doc/manuscript/response_to_reviews.Rmd

# Response to Reviewer(s)' Comments to Author:

## Reviewer: 1

The research is overall a worthwhile topic regarding joint effects of climate, tree size, and year on annual tree growth derived from tree-ring records of ten globally distributed forests.
I found the paper to be well-written and I appreciate the work you have accomplished. I consider the approach combining simultaneously the different factors controlling tree growth (non-linear effects of primary climate drivers, reconstructed tree diameter (DBH), and calendar year) highly appropriate for exploring species and stand responses to global changes in different biomes.
The study is well performed and the results and their interpretations are sound. The results are convincing and tree-ring data are nicely discussed in the context of global changes. However, given the complexity of the results, it would be very useful to report a summary picture of the main results by biome/ecoregions and level of naturalness of the studied stand (old-growth, seminatural forest ecosystems or secondary forests, reforestation/afforestation if present).

**We have added a summary figure (new Fig. 3) that we believe provides a very helpful summary. Thanks for the suggestion!**

@@

Figure 4: This is really interesting. It looks like there is often little effect of climate on the growth of small trees, which seems expected but I’m not sure has been shown like this before.

**Thank you!**

Line 545-547: Agree that this is useful to help with prediction, but what is missing from the article is a more clear idea of how this could/would be done. What emerges here is that trees and forests are complicated. There seem to be some general patterns emerge, but for each of these there are always exceptions. How do you propose synthesizing all of this to actually improve predictions?

**We agree that forests are complex, and this method both reveals and allows us to start to distill some of that complexity. We have added a summary figure (new Fig. 3) to help to highlight some of the general patterns. There is obviously a lot more work to be done on this theme that is beyond the scope of the current analysis.**

### 2021-08-04 doc/manuscript/ForestGEO_treerings.Rmd
summary fig caption & response to reviews
#136
modest rearrangement to create "Full model results overview" subsection in results

##GLS
GLS_BCNM = "S10"
GLS_SCBI = "S11"
GLS_SCBI_year = "S12"
GLS_LT = "S13"

@@

GLS_BCNM_legend = "Figure S10. Best GLS models including climate and DBH for BCNM"
GLS_SCBI_legend = "Figure S11. Best GLS models including climate and DBH for SCBI"
GLS_SCBI_year_legend = "Figure S12. Best GLS models including climate, DBH, and year for SCBI"
GLS_LT_legend = "Figure S13. Best GLS models including climate and DBH for LT"

@@

Fig. `r SummaryFigure`

```{r , eval=FALSE, echo=FALSE}
# This draw.io diagram that can be edited online here: https://app.diagrams.net/#HEcoClimLab%2FForestGEO-tree-rings%2Fmaster%2Fdoc%2Fmanuscript%2Ftables_figures%2FSummaryFigure.drawio
```
![**Figure 3 | Summary of top models for ring width (RW) as a function of climate (water and temperature variables), DBH, and calendar year**. Arrows approximate reponses detailed in Figs. 3, 5, and S16. Each symbol indicates one species, and species are ordered alphabetically within each site. Interactive effects of climate and DBH are not shown.](tables_figures/SummaryFigure.png)

@@

![**Figure 4 | Species-level responses of RW to climwin-selected variables in precipitation and temperature variable groups.** Primary climate drivers are coded on the x-axes as the climate variable abbreviation followed by the range of months (p=previous year, c=current year) over which it is most influential. For each species (color-coded as in Fig. 6), relationships are plotted if included in the top model. For each relationship shown, other terms in the model are held constant at their medians. Best-fit polynomials are plotted with solid lines when both first- and second-order terms are significant (t-test’s p-value <0.05), dash-dotted lines when only one term is significant, and dotted lines when neither is significant. Transparent ribbons indicate 95% confidence intervals. Vertical grey lines indicate the long-term mean for the climate driver over the analysis period; shading indicates 1 SD.](tables_figures/pre_temp_groups.png)

@@

When a precipitation variable, a temperature variable, and DBH were all included as candidate variables in the GLS models, typically all three were selected for the top model, regardless of the growth metric used.
Climate responses were generally similar to those described above for models without a DBH term, although some of the weaker climate responses were not consistently included in top models (e.g., $T_{min}$ responses at BCNM; Figs. `r clim_sens`, `r GLS_BCNM`).
In general, DBH explained more variation in growth rates than did climate, but its relative importance varied across growth metrics and sites (Figs. `r clim_sens`, `r size`).
The relative importance of DBH tended to be least for RW, intermediate for BAI, and highest for $\Delta$AGB (e.g., at SCBI; Fig. `r GLS_SCBI`), but exceptions occurred when RW decreased steeply with DBH (e.g., at LT, Fig. `r GLS_LT`).

@@

Growth rate -- whether measured as RW, BAI, or $\Delta$AGB -- varied with DBH for most species at all sites (Fig. `r size`).
These relationships were nonlinear for the majority of site-species combinations (`r round(n_spst_RW_DBH_bm/ n_spst_DBH_analyzed*100)` - `r round(n_spst_dAGB_DBH_best_model/ n_spst_DBH_analyzed*100)`% depending on growth metric; Fig. `r size`).

@@

![**Figure 6 | Growth sensitivity to DBH: (a) RW, (b) BAI, (c) $\Delta$AGB.** Relationships for species are plotted when included in the top model. Other terms in the model are held constant at their medians. Best-fit polynomials are plotted with solid lines when both first- and second-order terms are significant (t-test’s p-value <0.05), dash-dotted lines when only one term is significant, and dotted lines when neither is significant.  Transparent ribbons indicate 95% confidence intervals.](tables_figures/DBH_responses.png)

@@

![**Figure 7 | Effect of year, when included in the best model, on BAI.** For each species (all listed), relationships are plotted if the year effect could be analyzed and was included in the top model. Other terms in the model are held constant at their medians. Best-fit polynomials are plotted with solid lines when both first- and second-order terms are significant (t-test’s p-value <0.05), dash-dotted lines when only one term is significant, and dotted lines when neither is significant.  Transparent ribbons indicate 95% confidence intervals. ](tables_figures/Year_responses_BAI_only.png)

@@

`r GLS_LT_legend`

>>>>

##GLS
GLS_BCNM = "S10"
GLS_SCBI = "S11"
GLS_SCBI_year = "S12"
GLS_LT_year = "S13"

@@

GLS_BCNM_legend = "Figure S10. Best GLS models including climate and DBH for BCNM"
GLS_SCBI_legend = "Figure S11. Best GLS models including climate and DBH for SCBI"
GLS_SCBI_year_legend = "Figure S12. Best GLS models including climate, DBH, and year for SCBI"
GLS_LT_year_legend = "Figure S13. Best GLS models including climate and DBH for LT"

@@

When a precipitation variable, a temperature variable, DBH, and calendar year were all included as candidate variables in the GLS models, the most common outcome was that all four were included in the top model and had statistically significant effects (p $\le$ 0.05), regardless of the growth metric used (Fig. `r SummaryFigure`).
In general, DBH and calendar year explained more variation in growth rates than did climate, but their relative importance varied across growth metrics and sites (e.g., Figs. `r GLS_BCNM` - `r GLS_LT_year`).
Climate responses were generally similar regardless of the other variables included, although some of the weaker climate responses were not consistently included in top models (e.g., $T_{min}$ responses at BCNM; Figs. `r clim_sens`, `r GLS_BCNM`).
In contrast, effects of DBH and year often interacted such that the shape of the DBH response curve or its inclusion in the top model were frequently modified by the inclusion of calendar year (Fig. `r SummaryFigure`).

```{r , eval=FALSE, echo=FALSE}
# This draw.io diagram that can be edited online here: https://app.diagrams.net/#HEcoClimLab%2FForestGEO-tree-rings%2Fmaster%2Fdoc%2Fmanuscript%2Ftables_figures%2FSummaryFigure.drawio
```
![**Figure 3 | Summary of top models for ring width (RW) as a function of climate (water and temperature variables), DBH, and calendar year**. Arrows approximate reponses detailed in Figs. 3, 5, and S16. Each symbol indicates one species, and species are ordered alphabetically within each site. Overlapping solid and dotted arrows for the same species indicate that the effect changed when *Year* was included in the model, with dotted arrows indicating the response without *Year*. Interactive effects of climate and DBH are not shown.](tables_figures/SummaryFigure.png)

@@

![**Figure 4 | Species-level responses of RW to climwin-selected variables in precipitation and temperature variable groups.** Models presented here include only climate variables as fixed effects. Primary climate drivers are coded on the x-axes as the climate variable abbreviation followed by the range of months (p=previous year, c=current year) over which it is most influential. For each species (color-coded as in Fig. 6), relationships are plotted if included in the top model. For each relationship shown, other terms in the model are held constant at their medians. Best-fit polynomials are plotted with solid lines when both first- and second-order terms are significant (t-test’s p-value <0.05), dash-dotted lines when only one term is significant, and dotted lines when neither is significant. Transparent ribbons indicate 95% confidence intervals. Vertical grey lines indicate the long-term mean for the climate driver over the analysis period; shading indicates 1 SD.](tables_figures/pre_temp_groups.png)

@@

@@

Growth rate -- whether measured as RW, BAI, or $\Delta$AGB -- varied with DBH for most species at all sites (Fig. `r size`).
Because the effects of calendar year could not be evaluated for all species (Fig. `r SummaryFigure`), the DBH trends described here refer to models without year. 
Relationships between growth rate and DBH were nonlinear for the majority of site-species combinations (`r round(n_spst_RW_DBH_bm/ n_spst_DBH_analyzed*100)` - `r round(n_spst_dAGB_DBH_best_model/ n_spst_DBH_analyzed*100)`% depending on growth metric; Fig. `r size`).

@@

![**Figure 6 | Growth sensitivity to DBH: (a) RW, (b) BAI, (c) $\Delta$AGB.** Models presented here include climate variables and DBH as fixed effects. Relationships for species are plotted when included in the top model. Other terms in the model are held constant at their medians. Best-fit polynomials are plotted with solid lines when both first- and second-order terms are significant (t-test’s p-value <0.05), dash-dotted lines when only one term is significant, and dotted lines when neither is significant.  Transparent ribbons indicate 95% confidence intervals.](tables_figures/DBH_responses.png)

@@

![**Figure 7 | Effect of year, when included in the best model, on BAI.**  Models presented here include climate variables, DBH, and year as fixed effects. For each species (all listed), relationships are plotted if the year effect could be analyzed and was included in the top model. Other terms in the model are held constant at their medians. Best-fit polynomials are plotted with solid lines when both first- and second-order terms are significant (t-test’s p-value <0.05), dash-dotted lines when only one term is significant, and dotted lines when neither is significant.  Transparent ribbons indicate 95% confidence intervals. ](tables_figures/Year_responses_BAI_only.png)

@@

`r GLS_LT_year_legend`
